### 2019-05-09 manuscript/abstract.tex
- Submission #1 final changes - added cover letter and manuscript proofs

% E: Here's the previous abstract.  The new one below is aligned more with the current Introduction and Results.
%
% The effect of polyploidy in diversification remains a contentious issue. On the one hand,  recent studies  that found that polyploids have slower speciation rates and higher extinction rates than diploids left scientist wondering if polyploidy is truly an evolutionary dead-end. On the other hand, botanist have found strong molecular support of multiple polyploidy events at the root of highly diverse clades which challenges the evolutionary dead-end conclusions reached by modeling approaches. We re-investigate the role of polyploidy in speciation and extinction from a new modeling perspective considering that patterns found in diversification models can be misleading and incorrectly attributed to polyploidy when other observed and unobserved plant traits are responsible of shaping diversification.  Using  statistically robust comparative phylogenetic approaches, we show that it is possible to  detect whether  the contribution of polyploidy to speciation and extinction is significant  under the presence of  other potential traits also affect diversification. We use the phylogeny, polyploidy, and breeding system data of 595 Solanaceae species to understand the contribution of polyploidy to diversification. We ask if Solanaceae polyploids are evolutionary dead-ends, and whether breeding system or some other unobserved traits are responsible of the patterns of diversification observed in the phylogeny.

% E:  Help making this less bland?  Except for the first sentence, which may be too grandiose?
If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be understood as consequences of selection at high levels of the biological hierarchy.
Identifying traits associated with diversification rate differences is tricky, though, because of the many traits available to consider and the statistical challenge of testing for associations from phylogenetic data.
Two traits that have been repeatedly suggested as drivers of differential diversification are whether a lineage is diploid or polyploid, and whether it is self-incompatible or self-compatible.
We investigate the role of each of these traits, and their interaction, on speciation and extinction rates in Solanaceae.
We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.
These results are largely robust to assumptions about whether diploidization occurs.
Finally, we show that allowing for state-dependent diversification affects conclusions about the relative contribution of different evolutionary pathways to self-compatible polyploids.

>>>>

If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be understood as consequences of selection at high levels of the biological hierarchy.
Identifying traits associated with diversification rate differences is tricky, however, because of the wide variety of traits to consider and the statistical challenges of testing for associations from comparative phylogenetic data.
Ploidy (diploid vs.\ polyploid states) and breeding system (self-incompatible vs.\ self-compatible states) have been repeatedly suggested as possibly drivers of differential diversification.
We investigate the roles of these traits, including their interaction, on speciation and extinction rates in Solanaceae.
We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.
These results are largely robust to allowing for diploidization.
Finally, we show that the lower rate of diversification for self-compatible diploids makes it less likely for that state to lie along the evolutionary pathway from self-incompatible diploids to polyploids.

### 2019-05-09 manuscript/discussion.tex

Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids when exploring across multiple clades of the angiosperm phylogeny \citet{mayrose_2011, mayrose_2015}. 
We obtain a broadly similar outcome, when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states, and breeding systems (\cref{figure:netdivall}A and B). 
The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids.
Simultaneous consideration of breeding system better explains the variation of the inferred net diversification rates among ploidy states. 
SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.
Among SC species, diploids and polyploids are associated with similar diversification rates (\cref{figure:netdivall}E).
These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration. 
The finding is unique, and implications may be of wide interest, because it evaluates the relative influence of these two exceptionally interesting traits in a comparative phylogenetic framework.

@@

% qualifying paragraph for pathways
We implicitly modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016,freyman_2017}.
Models that fail to consider such cladogenetic changes may make misleading estimates of anagenetic transition rates.
In the only available data bearing on this problem, \citet{goldberg_2012} found that extending the *SSE framework to allow cladogenetic changes did not substantially affect inference of net diversification rates.
Nevertheless, a distinct possibility remains that allowing for cladogenetic transitions could tip the scales further in favor of the one-step pathway to SC polyploidy (CP), but this remains to be tested in future work. %B Will's comment alluded to, but not directly. 
Specifically, the multi/hidden state SSE-based framework we employed here could be extended to incorporate cladogenetic transitions. %B: We could also list %W+%R idea above, for stochastic mapping?
Another useful line of inquiry would examine the path-dependent diversification rates, asking whether diversification rates of polyploid SC lineages differ depending on whether they were generated via a one- or two-step pathway. %B: could elaborate if you see necessary. TODO
%
% examine SC-only families
% no causal correlation in SSI systems
% paths to dioecy
% path-dependent net-diversification
%
%B: also missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.
%
%B: And, are diversification rates of polyploid SC lineages different if one or other path was taken? % E: This is a great question.  We don't answer it with this round of analyses.  But discuss?
% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidden values, corresponding to the influx from the two pathways.  Those two states, CP_X and CP_Y, could have different lambda and mu.
%
%W: 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
%    the rest of the discussion, so modify or discard as needed:
% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012

@@

On the other hand, there is little evidence for the occurrence of re-diploidization within Solanaceae, since the origin of the family---the time frame we considered.
First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46-53 My \citealt{sarkinen_2013}). 
The best-supported, most-recent WGD \citet{tomato2012} is older than the lineage under consideration in this study.
At best then, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD, but likely before the origin of the family.
Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
Despite the limitations of assignment of ploidy levels, we found some support for models that include parameter $\delta$, meaning that further study may be warranted, especially in other groups.

Regardless of whether genome polypoidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that our present understanding of the evolution of ploidy and breeding systems is incomplete.
An increasingly forceful weight of evidence seems to support WGDs in the ancestry of many angiosperms. %cite
It is then easily inferred that many species have undergone diploidization, alongside genome size and chromosome number reductions. %cite
Inference of such a common genome upheaval in eudicot history seems to clash with the data indicating that a homologous mechanism of SI, which breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002,mcclure_2009}, has been present continually in many families, including Solanaceae \citep{igic_2006}. 
Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in this and other families.
As well, chromosome numbers and genome comparisons within the family (esp. `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca. 10 inversions and 11 translocations (likely underestimated).
Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
In either case, all of the approaches make ample assumptions, and it seems that at least some of them will necessitate deep revisions.
Our study, for example, made numerous assumptions in coding trait states (e.g. what constitutes a polyploid or diploid, SI or SC species), in models used for analyses (e.g. irreversibility of SI loss underlain by trans-specific orthologous S-loci), and many more, which could each mislead inference. %B: -> %R: Rosana, take a look at this.
% no discussion of magnitude of \delta vs \rho!
% R- magnitude is not easy to discuss about, what we have are estimates and credible intervals which are completely sample size dependent. With a different sample "magnitude" can change.
% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.

@@

%B: Rosana, read this carefully, please! I think it may need review:
%B: I changed nothing here, except: (1) a citealt<-citet to avoid double parentheses. (2) sp. disscussed -> discussed, (3) "If that is the case ON a system" -> "... IN a system"? (4) "hight type I errors" -> high, type I errors [comma], (5) line break in penultimate sentence.
Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in  \citealt{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification. 
We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.  
If that is the case in a study system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process. 
When considering a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood. 
Since MuSSE might also suffer from high type I errors, considering more heterogeneity using a hidden state model on top of the multi-state diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
% R- new lines
Investigating the process of diversification linked to traits can be done via thorough statistical inferences that carefully evaluate multiple evolutionary histories. 
Background heterogeneity in lineage of diversification is the rule, rather than the exception. 
Inferences that question whether other complex trait interactions are hidden in this background heterogeneity are necessary because they critically contribute to the reconstruction and understanding of diversification.

>>>>

Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across multiple clades of the angiosperm phylogeny \citep{mayrose_2011, mayrose_2015}. 
We obtain a broadly similar outcome when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states and breeding systems (\cref{figure:netdivall}A and B). 
The addition of a hidden trait blurs differences in diversification rates between diploids and polyploids.
Simultaneous consideration of breeding system better explains the variation of the inferred net diversification rates among ploidy states. 
SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.
Among SC species, diploids and polyploids are associated with similar diversification rates (\cref{figure:netdivall}E).
These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration. 
The finding is unique, and implications may be of wide interest, because it evaluates the relative influence of these two exceptionally interesting traits in a comparative phylogenetic framework.

@@

% qualifying paragraph for pathways
We modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016,freyman_2017}.
Models that fail to consider such cladogenetic changes may make misleading estimates of anagenetic transition rates.
In the only available data bearing on this problem, \citet{goldberg_2012} found that extending the *SSE framework to allow cladogenetic changes did not substantially affect inference of net diversification rates.
Nevertheless, a distinct possibility remains that allowing for cladogenetic transitions could tip the scales further in favor of the one-step pathway to SC polyploidy (CP), but this remains to be tested in future work. %B Will's comment alluded to, but not directly. 
Specifically, the multi/hidden state SSE-based framework we employed here could be extended to incorporate cladogenetic transitions. %B: We could also list %W+%R idea above, for stochastic mapping?
Another useful line of inquiry would examine the path-dependent diversification rates, asking whether diversification rates of polyploid SC lineages differ depending on whether they were generated via a one- or two-step pathway. %B: could elaborate if you see necessary. TODO
%
% examine SC-only families
% no causal correlation in SSI systems
% paths to dioecy
% path-dependent net-diversification
%
%B: also missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.
%
%B: And, are diversification rates of polyploid SC lineages different if one or other path was taken? % E: This is a great question.  We don't answer it with this round of analyses.  But discuss?
% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidden values, corresponding to the influx from the two pathways.  Those two states, CP_X and CP_Y, could have different lambda and mu.
%
%W: 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
%    the rest of the discussion, so modify or discard as needed:
% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012

@@

On the other hand, there is little evidence for the occurrence of re-diploidization within Solanaceae, since the origin of the family---the time frame we considered.
First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46-53 My \citealt{sarkinen_2013}). 
The best-supported, most-recent WGD \citep{tomato2012} is older than the lineage under consideration in this study.
At best then, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD, but likely before the origin of the family.
Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
Despite the limitations of assignment of ploidy levels, we found some support for models that include parameter $\delta$, meaning that further study may be warranted, especially in other groups.

Regardless of whether genome polypoidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that our present understanding of the evolution of ploidy and breeding systems is incomplete.
An increasingly forceful weight of evidence seems to support WGDs in the ancestry of many angiosperms. %cite
It is then easily inferred that many species have undergone diploidization, alongside genome size and chromosome number reductions. %cite
Inference of such a common genome upheaval in eudicot history seems to clash with the data indicating that a homologous mechanism of SI, which breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002,mcclure_2009}, has been present continually in many families, including Solanaceae \citep{igic_2006}. 
Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in this and other families.
As well, chromosome numbers and genome comparisons within the family (esp. `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations (likely underestimated).
Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
In either case, all of the approaches make ample assumptions, and it seems that at least some of them will necessitate deep revisions.
Our study, for example, made numerous assumptions in coding trait states (\eg what constitutes a polyploid or diploid, SI or SC species), in models used for analyses (\eg irreversibility of SI loss underlain by trans-specific orthologous S-loci), and many more, which could each mislead inference. %B: -> %R: Rosana, take a look at this.
% no discussion of magnitude of \delta vs \rho!
% R- magnitude is not easy to discuss about, what we have are estimates and credible intervals which are completely sample size dependent. With a different sample "magnitude" can change.
% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.

@@

%B: Rosana, read this carefully, please! I think it may need review:
%B: I changed nothing here, except: (1) a citealt<-citet to avoid double parentheses. (2) sp. disscussed -> discussed, (3) "If that is the case ON a system" -> "... IN a system"? (4) "hight type I errors" -> high, type I errors [comma], (5) line break in penultimate sentence.
Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in  \citealt{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focal trait first used to model diversification. 
We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.  
If that is the case in a study system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process.  % E: I would say not necessarily.  There could be other non-trait weirdnesses.
When considering a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood. 
Since MuSSE might also suffer from high type I errors, considering more heterogeneity using a hidden state model on top of the multi-state diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
% R- new lines
Investigating the process of diversification linked to traits can be done via thorough statistical inferences that carefully evaluate multiple evolutionary histories. 
Background heterogeneity in lineage of diversification is the rule, rather than the exception. 
Inferences that question whether other complex trait interactions are hidden in this background heterogeneity are necessary because they critically contribute to the reconstruction and understanding of diversification.

### 2019-05-09 manuscript/methods.tex

Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
After a first full round of curation via CCDBcurator, we contrasted results by hand and corrected chromosomes where CCDBcurator's output was not able to adequately correct records.
Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011} and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
%B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is. R- I added more (it only took 4months to hand-curate 14,000 entries
By comparing three different sources of information, we were able to code taxa as diploid (D) or polyploid (P).
For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on SI/SC classification.
For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid. 
Additionally, because of the absence of SI polyploids (discussed further below), species known to be SI could be scored as diploid.
Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.

Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
Many species could unambiguously be coded as either I or C \citep{raduski_2012}.
Following previous work, we coded any species with a functional SI system as I, even if SC or dioecy was also reported.
Dioecious species without functional SI were coded as C.
% E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
%R- Okay- I was getting confused as well thanks for clarifying

The Supplementary Information contains citations for the numerous original data sources. % todo: say how many refs?
Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. 
Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
Following the reasoning outlined in \citet{robertson_2011}, we closely examined the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population.
The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
Because of the resulting absence of SI and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
%B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.

Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
Binary or three-state classification of ploidy and breeding system for the 595 taxa is summarized in \cref{figure:stateclassifications}.
We retained all of these species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
A total of 405 taxa without any information about breeding system or ploidy were excluded.
%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
% R-This is very important but I have removed it because I think it is a full on paper by its own
%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.

@@

Both of our methods are based on a propagation matrix that describes flow from ID to CP, as in \citet{robertson_2011}.
We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
%
Second, we consider not only transitions between states but also diversification within each state.
State-dependent diversification can change the relative contributions of the two pathways.
In particular, if the net diversification rate is small for CD, the two-step pathway will contribute relatively less.
We therefore include the difference between speciation and extinction along the diagonal elements of the propogation matrix.
As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population. % R- here a reference would be nice

% "one-step predictions for P_I and P_C over time"
% E & R todo later: for the case with net div, write out the propagation matrix and equations for pathway contributions
%B: I corrected all instances of "propogation" in text and comments

>>>>

Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
After a first full round of curation via CCDBcurator, we contrasted results by hand and corrected chromosomes where CCDBcurator's output was not able to adequately correct records.
Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011} and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
%B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is. R- I added more (it only took 4months to hand-curate 14,000 entries
By comparing three different sources of information, we were able to code taxa as diploid (D) or polyploid (P).
For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on SI/SC classification.
For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid. 
Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.

Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
Most species could unambiguously be coded as either I or C \citep{raduski_2012}.
Following previous work, we coded any species with a functional SI system as I, even if SC or dioecy was also reported.
Dioecious species without functional SI were coded as C.
% E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
%R- Okay- I was getting confused as well thanks for clarifying

The Supplementary Information contains citations for the numerous original data sources. % todo: say how many refs?
Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. 
Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
Following the reasoning outlined in \citet{robertson_2011}, we closely examined the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population.
The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
Because of the resulting absence of polyploid SI populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
%B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.

Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
Of these, 290 had information for both states.
The number of species in each combination of states is summarized in \cref{figure:stateclassifications}.
We retained all 595 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
A total of 405 taxa without any information about breeding system or ploidy were excluded.
%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
% R-This is very important but I have removed it because I think it is a full on paper by its own
%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.

@@

Both of our methods are based on a propagation matrix that describes flow from ID to CP, as in \citet{robertson_2011}.
We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
%
Second, we consider not only transitions between states but also diversification within each state.
State-dependent diversification can change the relative contributions of the two pathways.
In particular, if the net diversification rate is small for CD, the two-step pathway will contribute relatively less.
We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population. % R- here a reference would be nice

% "one-step predictions for P_I and P_C over time"
% E & R todo later: for the case with net div, write out the propagation matrix and equations for pathway contributions
